---
title: "Exploring Musical Trends: Analyzing Song Durations, Keys, and Energy Levels"
subtitle: "A Comparative Study of A Boogie Wit Da Hoodie, Pay Money To My Pain, and Radiohead Using Spotify Metadata"
author: 
  - Lexun Yu
  - Colin Sihan Yang
  - Siddharth Gowda
  - Yuxuan Wei
date: today
date-format: long
abstract: "This paper analyzes metadata for songs by A Boogie Wit Da Hoodie, Pay Money To My Pain (PTP), and Radiohead, focusing on trends in song durations, key usage, and energy levels. A Boogie’s recent albums feature shorter songs, reflecting modern media influences, while PTP’s 2014 album shows wide duration variability, likely tied to emotional tributes to their late vocalist. Radiohead maintains consistent song durations, emphasizing artistic experimentation over trends. Key analysis reveals A Boogie favors minor keys, with consistent energy levels across key types."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

library(spotifyr)
library(styler)
library(tidyverse)
library(tinytex)
library(usethis)
library(lintr)
library(knitr)
library(here)
library(ggplot2)
library(janitor)
```

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

PTP <- readRDS(here("data/01-raw_data/PTP.rds"))
radiohead <- readRDS(here("data/01-raw_data/radiohead.rds"))
boogie <- readRDS(here("data/01-raw_data/boogie.rds"))
clean_boogie_data <- read_csv(here("data/02-cleaned-data/clean_boogie_data.csv"))
```

# Introduction
This paper download data of three artists "A Boogie Wit da Hoodie", "Radiohead" and "Pay money To my Pain" from Spotify API, and visualize the metadata obtained. First, we compare the length of released songs between the three artists in @sec-date-duration. Second, we types of keys in "A Boogie Wit da Hoodie", and visualize the Major and Minor keys in their songs in @sec-key.

The analysis shows A Boogie Wit Da Hoodie’s recent albums have shorter songs, PTP’s 2014 album features a wide range of durations reflecting tributes to their late vocalist, and Radiohead maintains consistent durations. A Boogie favors minor keys, with contrasting usage of C# in major and minor keys, and energy levels are consistent across key types.

This paper is written with the help of R [@citeR], and the following packages: spotifyr [@spotifyr], styler [@styler], tidyverse [@tidyverse], tinytex [@tinytex], usethis [@usethis], lintr [@lintr], here [@here], ggplot2 [@ggplot2], janitor [@janitor] and knitr [@knitr]. The data is downloaded from Spotify API [@spotify_api].


# Data
## Release Date and Duration {#sec-date-duration}
```{r}
#| label: fig-dateduration
#| fig-cap: Comparison of Song Durations Across Album Release Dates for Different Artists 
#| fig-subcap: ["Boogie’s More Recent Albums Feature Shorter Songs","Ptp’s 2014 Album Shows a Wide Range of Song Durations", "Radiohead Maintains Consistent Song Durations Over Time"] 
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 3

boogie <- as_tibble(boogie)

boogie |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms /60000,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (min)"
  )

PTP |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms / 60000,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (min)"
  )

radiohead |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms / 60000,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (min)"
  )

```

These three boxplots in in @fig-dateduration display the relationship between album release dates and song durations (in minutes) for A Boogie Wit Da Hoodie, Pay Money to my Pain(PTP), and Radiohead. In @fig-dateduration-1, for A Boogie Wit Da Hoodie, the data reveals a trend where more recent albums feature shorter songs, possibly reflecting the influence of short-form video platforms like TikTok, which encourage brevity in content.

For Pay Money To My Pain (PTP) in @fig-dateduration-2, the 2014 album stands out as significant. This album was released shortly after the death of the band's vocalist, K, in the same year. The boxplot shows a wide range of song durations for this album, likely reflecting the use of many vocal recordings made by K before his passing. This range may also capture the emotional intensity and variation in the music as the band paid tribute to their late vocalist, making it a deeply personal and diverse work.

In contrast, in @fig-dateduration-3, Radiohead exhibits less of this trend, maintaining relatively consistent song durations over time. This might indicate a different creative approach, as Radiohead’s music tends to focus more on experimentation and artistry, rather than following trends in popular media. Together, these boxplots suggest that while some artists adapt to new trends, others continue to focus on traditional album structures.

## Keys in the songs of "A Boogie Wit da Hoodie" {#sec-key}
In this section we discuss the keys "A Boogie Wit da Hoodie" uses the most, and find the most frequently used Major and Minor keys used by the artist. 

```{r}
#| label: fig-majorvminor
#| fig-cap: "Boogie Uses Major Keys Around C# The Most"
#| echo: false
#| warning: false
#| message: false

clean_boogie_data %>% ggplot(aes(mode_name)) +
  geom_bar() + theme_minimal() + labs(x = "Key Type (Major or Minor)", y = "Number of Songs")


```
Based on @fig-majorvminor, Boogie With Da Hoodie has released more minor key songs than major key songs. 


```{r}
#| label: fig-allkeycounts
#| fig-cap: "Boogie Uses Minor Keys Around C# The Least and the F and G Minor Keys Often"
#| fig-subcap: ["Major", "Minor"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

clean_boogie_data %>% filter(mode_name == "major") %>% ggplot(aes(key_name)) +
  geom_bar() + theme_minimal() + labs(x = "Key Name", y = "Number of Songs in Key")

clean_boogie_data %>% filter(mode_name == "minor") %>% ggplot(aes(key_name)) +
  geom_bar() + theme_minimal() + labs(x = "Key Name", y = "Number of Songs in Key")

```

Based on @fig-allkeycounts, in major key, Boogie uses major keys around C# the most. However for minor keys this is the exact opposite, meaning he uses minor keys around C# the least. Instead he uses the F and G minor keys often.

```{r}
#| label: fig-energy
#| fig-cap: "Energy Level Distribution For Boogie Songs is Normal and Sysmetric"
#| fig-subcap: ["Major", "Minor"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

# For Discussion say they based on this hist there doesn't seem to be a difference
# in the energy levels of a song given the key

clean_boogie_data %>% filter(mode_name == "major") %>% ggplot(aes(energy)) +
  geom_histogram(binwidth = 0.03) + theme_minimal() + 
  labs(x = "Energy", y = "Number of Songs")

clean_boogie_data %>% filter(mode_name == "minor") %>% ggplot(aes(energy)) +
  geom_histogram(binwidth = 0.03) + theme_minimal() + 
  labs(x = "Energy", y = "Number of Songs")

```

Based on @fig-energy, both the distribution of energy level per songs for major and minor keys appear to be approximatley normal and sysmetric. Moreover, both histograms show that Boggie's songs energy levels are usually around 0.6.

\newpage

# Bibliography