---
title: "Analysis of ACS 2022"
author: 
  - Lexun Yu
  - Colin Sihan Yang
  - Siddharth Gowda
date: today
date-format: long
abstract: "Use the Spotify API to analyze data about ... "
format: pdf
number-sections: true
bibliography: references.bib
---


# Introduction



# Data

```{r}
clean_boogie_data <- read_csv("../data/clean_data.csv")

```


# Results
```{r}
#| label: fig-majorvminor
#| fig-cap: "Number of Songs In Major Key"
#| echo: false
#| warning: false
#| message: false

clean_boogie_data %>% ggplot(aes(mode_name)) +
  geom_bar() + theme_minimal() + labs(x = "Key Type (Major or Minor)", y = "Number of Songs")


```
Based on @fig-majorvminor, Boogie With Da Hoodie has released more minor key songs than major key songs. 


```{r}
#| label: fig-allkeycounts
#| fig-cap: "Number of Songs Per Key"
#| fig-subcap: ["Major", "Minor"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

clean_boogie_data %>% filter(mode_name == "major") %>% ggplot(aes(key_name)) +
  geom_bar() + theme_minimal() + labs(x = "Key Name", y = "Number of Songs in Key")

clean_boogie_data %>% filter(mode_name == "minor") %>% ggplot(aes(key_name)) +
  geom_bar() + theme_minimal() + labs(x = "Key Name", y = "Number of Songs in Key")

```
Based on @fig-allkeycounts, in major key, Boogie uses major keys around C# the most. However for minor keys this is the exact opposite, meaning he uses minor keys around C# the least. Instead he uses the F and G minor keys often.

```{r}
#| label: fig-energy
#| fig-cap: "Energy Level Distribution For Boogie Songs"
#| fig-subcap: ["Major", "Minor"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

# For Discussion say they based on this hist there doesn't seem to be a difference
# in the energy levels of a song given the key

clean_boogie_data %>% filter(mode_name == "major") %>% ggplot(aes(energy)) +
  geom_histogram(binwidth = 0.03) + theme_minimal() + 
  labs(x = "Energy", y = "Number of Songs")

clean_boogie_data %>% filter(mode_name == "minor") %>% ggplot(aes(energy)) +
  geom_histogram(binwidth = 0.03) + theme_minimal() + 
  labs(x = "Energy", y = "Number of Songs")

```
Based on @fig-energy, both the distribution of energy level per songs for major and minor keys appear to be approximatley normal and sysmetric. Moreover, both histograms show that Boggie's songs energy levels are usually around 0.6.

