---
title: "Analysis of Hitsujibungaku's Spotify Data"
author: 
  - Lexun Yu
  - Colin Sihan Yang
  - Siddharth Gowda
  - Yuxuan Wei
date: today
date-format: long
abstract: "Use the Spotify API to analyze data about ... "
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

library(spotifyr)
library(styler)
library(tidyverse)
library(tinytex)
library(usethis)
library(lintr)
library(knitr)
library(here)
```

```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false

PTP <- readRDS(here("data/01-raw_data/PTP.rds"))
radiohead <- readRDS(here("data/01-raw_data/radiohead.rds"))
boogie <- readRDS(here("data/01-raw_data/boogie.rds"))
```

# Introduction
This paper download data of three artists "A Boogie Wit da Hoodie", "Radiohead" and "Pay money To my Pain" from Spotify API, and visualize the metadata obtained. First, we compare the length of released songs between the three artists. Second, we 

This paper is written with the help of R [@citeR], and the following packages: spotifyr [@spotifyr], styler [@styler], tidyverse [@tidyverse], tinytex [@tinytex], usethis [@usethis], lintr [@lintr], here [@here] and knitr [@knitr].


```{r}
boogie <- as_tibble(boogie)

boogie |>
  mutate(album_release_date = ymd(album_release_date)) |>
  ggplot(aes(
    x = album_release_date,
    y = duration_ms,
    group = album_release_date
  )) +
  geom_boxplot() +
  geom_jitter(alpha = 0.5, width = 0.3, height = 0) +
  theme_minimal() +
  labs(
    x = "Album release date",
    y = "Duration of song (ms)"
  )
```

# Bibliography